---
title: "Phone Number"
author: "Steve Myles"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction to phonenumber

This document describes the usage of the two functions in the `phonenumber`
package. Note that, though this package's functions convert back and forth from
numbers to English letters as on a telephone's keypad.  Note that the length of
the character string need not conform to any standard length telephone number.

### phonenumber's functions

* `letterToNumber`
* `numberToLetter`

### Telephone keypad

For purposes of this package, the mapping of numbers to letters on a telephone's
keypad are as follows:

* 2 corresponds to A, B, C
* 3 corresponds to D, E, F
* 4 corresponds to G, H, I
* 5 corresponds to J, K, L
* 6 corresponds to M, N, O
* 7 corresponds to P, Q, R, S
* 8 corresponds to T, U, V
* 9 corresponds to W, X, Y, Z
* 1 and 0 do not have corresponding letters

(I am aware that on some telephones, primarily older models, Q and Z are mapped
to 0 rather than to 7 and 9.)

```{r library, echo=FALSE}
library(phonenumber)
```

## `letterToNumber`

`letterToNumber` converts a string containing letters into the corresponding 
numbers on a telephone's keypad. For example, if the user wants to know what telephone number corresponds to "Texas:"

```{r letterToNumber ex1}
string <- "Texas"
letterToNumber(string)
```

The function is not limited to letters, though. Any numeric characters are
returned as is and any non-alphanumeric characters (including spaces) are 
converted to dashes (-) and 1 and 0 are returned as is:

```{r letterToNumber ex2, echo = TRUE}
string <- "Texas is #1"
letterToNumber(string)
```

## `numberToLetter`

`numberToLetter` converts a string containing numbers into the corresponding 
letters on a telephone's keypad. For example, if the user wants to know what 
possible character strings could be spelled by a sequence of numbers (e.g., 22):

```{r numberToLetter ex1}
string <- "22"
numberToLetter(string)
```

The function is not limited to numbers, though. Any alphabetic characters are
returned as is and any non-alphanumeric characters (including spaces) are 
converted to dashes (-) and 1 and 0 are returned as is:

```{r numberToLetter ex2, echo = TRUE}
string <- "806!"
numberToLetter(string)
```